To talk about graphs, we first have to agree on some notation: Let $V$ be a set and $E\subset V\times V$. The tuple $G=(V,E)$ is called a graph. Furthermore, we call an element $x\in V$ a node and a tuple 
$(x, y)\in E$ a directed edge from $x$ to $y$. Suppose there is a directed edge from $x$ to $y$, then we say that $x$ is a neighbor of $y$. 
In case $(x,y)\in E$ implies $(y,x)\in E$, we speak of an undirected graph.
In that case, we can think of elements in $E$ as unordered tuples $\{x,y\}$ instead of ordered ones. We still call $\{x,y\}$ an edge between $x$ and $y$.
For $y\in V$ we define the neighborhood $N_y$ of $y$ to be the set of all neighbors, i.e.
\begin{equation}
    \label{eq:def_neighbors}
    N_y\coloneq\left\{x\in V : (x,y)\in E\right\}\subset V.
\end{equation}
Furthermore, we assign to each node $x\in V$ a vector $v_x\in\R^n$ called node feature and to each edge $(x, y)\in E$ a vector $e_{x,y}\in\R^m$ called edge feature.
A graph together with all its node and edge features is called attributed graph. 

Roughly, a GNN takes an attributed graph as an input and manipulates the edge and node features in each step. 
More than that, a GNN can transform the structure of the graph itself, e.g. by introducing new nodes or edges. 
However, we will not go into detail about this possibility and stick to the simpler case of manipulating only node and edge-features.
Furthermore, we restrict ourselves to the case where the GNN does not alter the edge features.
\begin{SCfigure}[1][ht]
\centering
\label{fig:gnn}
\begin{tikzpicture}[
    node_r/.style={circle, draw=red!60, fill=red!5, very thick, minimum size=7mm},
    node_o/.style={circle, draw=orange!60, fill=orange!5, very thick, minimum size=7mm},
    node_g/.style={circle, draw=gray!60, fill=gray!5, very thick, minimum size=7mm},
    node/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm},
    mess/.style={rectangle, draw=green!60, fill=green!5, very thick, minimum size=7mm},
    ]
    %Nodes step 1
    \node[node_r] (centernode)                          {$v_c$};
    \node[node] (top)           [above=of centernode]   {$v_x$};
    \node[node_o] (bottom)        [below=of centernode]   {};
    \node[node] (right)         [right=of centernode]   {$v_y$};
    \node[node] (left)          [left=of centernode]    {$v_z$};
    \node[node_g] (llt)           [above left=of left]    {};
    \node[node_g] (llb)           [below left=of left]    {};
    \node[node_g] (rrt)           [above right=of right]    {};
    \node[node_g] (rrb)           [below right=of right]    {};

    %Nodes step 2
    \node[mess] (mess_top)           [below=of bottom]   {$m_x$};
    \node[node_r] (centernode_2)  [below=of mess_top]     {$v_c$};
    \node[mess] (mess_right)         [right=of centernode_2]   {$m_y$};
    \node[mess] (mess_left)          [left=of centernode_2]    {$m_z$};

    %Nodes step 3
    \node[node_r] (centernode_3)  [below=of centernode_2]     {$v_c$};
    \node[mess] (mess_final)      [left=of centernode_3]   {$m_{x+y+z}$};

    %Nodes step 4
    \node[node_r] (centernode_4)  [below=of centernode_3]     {$\widehat{v_c}$};
    
    %Lines step 1
    \draw[->, thick]  (top.south) -- node[anchor=west]{$e_{x,c}$} (centernode.north);
    \draw[->, thick] (left.east) -- node[anchor=south]{$e_{z,c}$} (centernode.west);
    \draw[->, thick] (right.west) -- node[anchor=north]{$e_{y,c}$} (centernode.east);
    \draw[->, thick, gray] (centernode.south) -- (bottom.north);
    \draw[->, thick, gray] (left.north west) -- (llt.south east);
    \draw[->, thick, gray] (left.south west) -- (llb.north east);
    \draw[->, thick, gray] (right.north east) -- (rrt.south west);
    \draw[->, thick, gray] (right.south east) -- (rrb.north west);

    %Lines step 2
    \draw[->, thick]  (mess_top.south) -- (centernode_2.north);
    \draw[->, thick] (mess_left.east) -- (centernode_2.west);
    \draw[->, thick] (mess_right.west) -- (centernode_2.east);

    %Lines step 3
    \draw[->, thick]  (mess_final.east) -- (centernode_3.west);

    \draw[->, very thick, gray] (-4,-0) -- node[anchor=east, black]{$\phi_{mes}$} (-4,-5.2);
    \draw[->, very thick, gray] (-4,-5.5) -- node[anchor=east, black]{$\phi_{agg}$} (-4,-7.2);
    \draw[->, very thick, gray] (-4,-7.5) -- node[anchor=east, black]{$\phi_{upd}$} (-4,-9.2);
    %\draw[->] (rightsquare.south) .. controls +(down:7mm) and +(right:7mm) .. (lowercircle.east);
\end{tikzpicture}
\caption{Illustration of the message passing procedure according to equation~\eqref{eq:mess_pass}. 
The neighbors of node $c$ (red) are the nodes $x,y,z$ (green). Attention: According to equation~\eqref{eq:def_neighbors} the orange node is not regarded as a neighbor of $c$ as the edge points in the wrong direction.
For each neighbor ($x,y,z$), $\phi_{mes}$ computes messages ($m_x,m_y,m_z$) which are sent to $c$.
Then, $\phi_{agg}$ aggregates these three messages and outputs one overall message $m_{x+y+z}$ that is sent to $c$. 
In the last step, $\phi_{upd}$ updates the node value $v_c$ to a new value $\widehat{v_c}$.  
}
\end{SCfigure}
Let us make these ideas a bit more rigorous (see figure~\ref{fig:gnn} for an illustration):  
Let $(V,E)$ be a graph with node features $\{v_x\in\R^n : x\in V\}$ and edge features $\{e_{x,y}\in\R^m:(x,y)\in E\}$.
For each $c\in V$ a new node feature $\widehat{v_c}$ is calculated according to the following rule
\begin{equation}
    \label{eq:mess_pass}
    \widehat{v_c}=\phi_{upd}\left(
        v_c, \phi_{agg}\left(\left\{
            \phi_{mes}\left(v_y,v_c, e_{y,c}\right):y\in N_c
            \right\}\right)
    \right),
\end{equation}
where $\phi_{upd},\phi_{agg},\phi_{mes}$ denote differentiable functions. These three functions are commonly interpreted as follows:
$\phi_{mes}$ takes as inputs the node value $v_c$ and the node value $v_y$ of one neighbor $y$ of $c$ as well as the value $e_{y,c}$ of the edge $(y,c)$. 
Depending on these inputs, it then computes a value, which can be thought of as a message originating from node $y$ which is sent to node $c$.
$\phi_{agg}$ collects all these messages to node $c$ and aggregates them in some way, so that the output can be thought of as one overall message to node $c$.
$\phi_{upd}$ takes this overall message as well as the value of node $c$ and computes how the value of node $c$ is altered.
Unsurprisingly, this scheme is called message passing layer. Given a specific problem that is required to be solved by a GNN, the challenge is to make appropriate choices
for $\phi_{upd},\phi_{agg},\phi_{mes}$ that suit the problem at hand. Furthermore, one has to decide how many iterations of the above procedure are suitable.